---
title: "Syllabus"
output: 
  html_document:
    includes:
      after_body: _includes/include_footer-no-disqus.html
      in_header: _includes/header_banner.html 
---

<!--
Update lines 199 and 220 if new class.
library(knitr)
rmarkdown::render_site("syllabus.Rmd")# build website
rmarkdown::render_site()# build website
-->


Instructor: Dr. Michael E. Colvin  
Office: Thompson 215  
Office phone: 662-325-3592  
Email: michael.colvin@msstate.edu  
Office hour(s): By appointment  
Lecture (M, W): THMANX 208, 8-8:50 am   
Lab (T): THMANX 313, 1:00-4:50 pm  
  
## Catalog description

Principles of fisheries management and methods for assessment and 
analysis of fish populations and aquatic habitats. Two hours lecture. 


## Textbook and course material

There is no required textbook for this class, supplementary PDFs will be 
provided as needed and available on the course Blackboard site 


##Course background

Contemporary fisheries managers require diverse skills and abilities to 
properly identify management actions and alternatives, as well as 
implement management programs. Additionally, developing monitoring 
programs for fish, habitat, and anglers is essential to evaluate 
management actions. Fisheries professionals also must: 


* Possess and understand the analytical tools necessary to interpret 
analyses of fisheries data and use the results to inform management and 
conservation decisions. 
* Be able to communicate with natural resource professionals and 
non-technical stakeholders 
* Collaborate with peers from multiple disciplines to successfully 
complete projects 
* Do more with less (e.g., reduced budgets, reduced personnel)
* Do more with more (e.g., big data, data loggers, mark recapture)
* Do your work safely, ethically, and under public scrutiny
* Interact with an increasingly informed public through social media, 
stakeholder meetings, and daily interactions 
* Allow public participation in decision making and management processes

## Course objectives

This course will provide you with a foundation of fisheries management 
and knowledge of commonly-used analytical techniques employed by 
fisheries managers to manage fisheries 



By the end of the course, you will be able to:

1. Define different fishery types, history and application
1. Compare and evaluate common practices used to manage and conserve 
fishes, fisheries and their habitats 
1. Understand population dynamics and the effects of harvest and 
propagation 
1. Calculate dynamic rate functions (i.e., recruitment, growth, and 
mortality) and population dynamics of fish populations 
1. Interpret and evaluate stock assessment techniques and their 
application to management and conservation 
1. Explain common population indices, including their assumptions, 
benefits, and limitations 
1. Understand contemporary management of natural resources

## Academic misconduct

Academic integrity is serious and all disciplinary action will be taken 
in the event of academic dishonesty. Mississippi State has an approved 
Honor Code that applies to all students. The code is as follows: "As a 
Mississippi State University student, I will conduct myself with honor 
and integrity at all times. I will not lie, cheat, or steal, nor will I 
accept the actions of those who do." Upon accepting admission to 
Mississippi State University, a student immediately assumes a commitment 
to uphold the Honor Code, to accept responsibility for learning, and to 
follow the philosophy and rules of the Honor Code. Student will be 
required to state their commitment on examinations, research papers, and 
other academic work. Ignorance of the rules does not exclude any member 
of the MSU community from the requirements or the processes of the Honor 
Code. For additional information, please visit: 
http://honorcode.msstate.edu/policy. 


## Cell Phone Policy

In accordance with Academic Operating Policy 10.08 (approved July 12, 
2005) and in order to limit classroom disruptions, as well as to protect 
against academic misconduct, the use by students of cell phones, 
messaging devices and other electronic devices is prohibited. In this 
class, students are required to put cell phone in the silent mode and 
stow in backpack while attending class and laboratory sessions. Cell 
phones may be used as calculators or clocks at any time during class and 
laboratory sessions. 


Professional Expectations for Students in the Wildlife, Fisheries and 
Aquaculture Program 


http://www.cfr.msstate.edu/wildlife/documents/professional_expectations.pdf

## Assessment of student progress

### 4313/6313 Students

Academic progress will be evaluated using a combination of exams, 
problem sets, a technical manuscript, in class presentation and 
participation. 


1. Exams (450 points).—Three exams will be administered during the 
semester. Exams will consist of short answer and essay questions, and 
will focus on your ability to summarize, evaluate, and interpret 
information using the basic knowledge you have gained in both the 
lecture and laboratory portions of the course. 
1. Laboratory problem sets (200 points).—During the course laboratory 
sessions you will develop or be given data set that you will analyze 
during lab and your own time. You will be required to conduct analyses 
using the tools and techniques provided during the course and labs as 
teams or individually. Time will be given for you to work on these 
problem sets during the labs, but additional time may be required 
outside of class for interpreting the results. 
1. Management brief (250 points total).—Contemporary fishery managers 
need to be able to communicate research and management actions to 
stakeholders using a variety of media. This may include Facebook, 
Twitter, articles in popular magazines (e.g., Field and Stream, 
Mississippi Outdoors), and orally in public meetings. During the 
semester you will select a published manuscript from an established 
fisheries journal and provide the following: 
    a. A management brief (175 points): a summary brief of the article that 
    would be appropriate for a popular outlet. There are no word limits for 
    this, with all writing you should only use the number of words needed to 
    accurately and succinctly convey your message. Grading of your brief 
    will be a two-step process that simulates “real-world” experiences you 
    will encounter in your career. First, you will submit a draft of the 
    brief that will be graded and returned to you with detailed comments. 
    After revision, a final version of the brief will be submitted and a 
    final grade will be assigned. 
    b. Presentation (75 points).—During the final week of laboratory you 
    will be required to present the details of the paper you selected as 
    brief 5 minute oral presentation. 
1. Class assignments and participation (100 points).—Part of being a 
fisheries manager is actively participating in discussions. The majority 
of fisheries management is working with people, not fish! You must be 
present in class and an actively participant in discussions and team 
work. 


### Summary of evaluations:

| Task                                	| Points 	|
|-------------------------------------	|:------:	|
| Exam I                              	|   125  	|
| Exam II                             	|   125  	|
| Exam III (Final)                    	|   200  	|
| Laboratory problem sets             	|   200  	|
| Management brief                    	|   250  	|
| Class assignments and participation 	|   100  	|
| Total                               	|  1000  	|			

### 6313 Students- Management model and description

Students enrolled in WFA6313 are required to develop a management model 
of their research system using STELLA. The model should include at 
least: 3 state variables, 2 hypotheses, and a decision utility. The 
final deliverable will be a report describing the model and a working 
STELLA model. Guidelines will be posted on the course Blackboard site. 


### Grading Scale

A= 895 to 1000  	B = 795 to 894		C = 695 to 794		D = 595 to 694
F = 0 to 594

### Late assignments and missing exam policies

All assignments must be turned in on time. Late assignments may be 
graded, but the maximum point value will be reduced by 5% for every day 
the assignment is late; however, assignments will not be graded if they 
are turned in after graded assignments are returned to the class. If for 
some reason there is a timing conflict associated with an exam, discuss 
these conflicts with me at least two weeks prior to the scheduled date. 

## Important Dates


```{r,echo=FALSE}
dat<-xlsx::read.xlsx("_includes/course-schedule.xlsx" ,sheetName="important-dates")
pander::pander(lapply(1:nrow(dat),function(x) as.character(dat$val)[x]))
```

## Students with Disabilities

Accommodations are collaborative efforts between students, faculty and Disability Support Services (DSS). Students with accommodations approved through DSS are responsible for contacting the faculty member in charge of the course prior to or during the first week of the term to discuss accommodations. Students who believe they are eligible for accommodations but who have not yet obtained approval through DSS should contact DSS immediately at (662) 325-3335.

## Title IX

MSU is committed to complying with Title IX, a federal law that prohibits discrimination, including violence and harassment, based on sex. This means that MSU' s educational programs and activities must be free from sex discrimination, sexual harassment, and other forms of sexual misconduct. If you or someone you know has experienced sex discrimination, sexual violence and/or harassment by any member of the University community, you are encouraged to report the conduct to MSU' s Director of Title IX/EEO Programs at 325-8124 or by e-mail to titleix@msstate.edu. Additional resources are available at:
* http://www.msstate.edu/web/security/title9-12.pdf, or 
* http://students.msstate.edu/sexualmisconduct/

  
## Tentative Lecture and Lab Schedule

Laboratory time subject to change with instructor availability; 
laboratory sessions may be used to make up for missed lectures due to 
weather and instructor availability.

```{r,echo=FALSE}
dat<-xlsx::read.xlsx("_includes/course-schedule.xlsx" ,sheetName="2017-F",
    stringsAsFactors=FALSE)
dat<-dat[,c(1:4)]
dat[is.na(dat)]<-""
knitr::kable(dat,
    col.names=c("Week","Monday (Class: 8-8:50)","Tuesday (Lab: 1-4:50)",
    "Wednesday (Class: 8-8:50)"))
```



